---
title: "Kaggle_Project2"
author: "Hollismc"
date: '2022-04-05'
output: html_document
---
##  Introduction  
  This dataset was gathered from Our World in Data from their article "Plastic Pollution". The authors inspiration for the data set was to show how much plastic we use per day.  This analysis will visualize how much plastic is produced and how much plastic we use and discard overall.  Then we ask what countries are above the mean in both mismanaged waste and per capita usage of plastic? While the authors don't exactly define what mismanaged waste is we surmise that it is the amount of plastic being dumped in the ocean.  They state that "More and more plastic is being generated every year and more of which is getting dumped into the ocean or is mishandled".

##  Libraries
```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(readr)
```
##  Download Data 
```{r}
plastpro <- read_csv("archive-2/global-plastics-production.csv")
mismanplast <- read_csv("archive-2/mismanaged-waste-global-total.csv")
plastwstpercap <- read_csv("archive-2/plastic-waste-per-capita.csv")
str(plastpro)
str(mismanplast)
str(plastwstpercap)
```
  We begin our analysis with a simple graph of global plastics production to show its dramatic rise from 1950 to 2015. We will then look at waste per capita and what the mismanagement of that waste means to our environment. We are limited to one year (2010) of data for plastic mismanagement and plastic waste per capita, but will be advised by the global plastics production's upward trajectory and trends for mismanagement and consumption.  We will identify those contries that contribute and mismanage more than the mean.  

```{r}
plastprodat <- plastpro %>% select(Year, `Global plastics production (million tonnes)`)
 plastplot <- ggplot(data = plastprodat, aes(plastprodat$Year, plastprodat$`Global plastics production (million tonnes)`)) +
            geom_point() +
            geom_abline() +
            labs(x="Year", y = "Global Plastics Production", title = "Global Plastics Production
                 In Million Tonnes")
```
##  Combine the data necessary into one data set          
```{r}
mismanplastdat <- mismanplast %>% select(Entity, `Mismanaged waste (% global total)`)
plastwstpercapdat  <- plastwstpercap %>% select(Entity, `Per capita plastic waste (kg/person/day)`)
plasmut <- right_join(mismanplastdat, plastwstpercapdat)
hist(plasmut$`Mismanaged waste (% global total)`)
hist(plasmut$`Per capita plastic waste (kg/person/day)`)

```
## filter out countries greater than the mean to identify the biggest offenders
```{r}
plwstpcdatmean <- mean(plastwstpercapdat$`Per capita plastic waste (kg/person/day)`)
mismanmean <- mean(mismanplastdat$`Mismanaged waste (% global total)`)
bigoff <- plasmut %>% filter(`Mismanaged waste (% global total)`> mean(`Mismanaged waste (% global total)`)) %>% filter(`Per capita plastic waste (kg/person/day)`> mean(`Per capita plastic waste (kg/person/day)`)) %>% select(Entity, `Mismanaged waste (% global total)`, `Per capita plastic waste (kg/person/day)`)
hist(bigoff$`Mismanaged waste (% global total)`)
hist(bigoff$`Per capita plastic waste (kg/person/day)`)
```
## Biggest Per Capita waste by Country in 2010
```{r}
bigoffplot <- ggplot(bigoff, aes(Entity, bigoff$`Per capita plastic waste (kg/person/day)`)) +
        geom_point() +
        labs(x="Country", y="Per Capita Plastic Waste (kg/person/day)", title = "Biggest Per Capita Waste by Country 2010")
```

## Biggest Mismanaged Waste by Country in 2010
```{r}
bigoffplot_2 <- ggplot(bigoff, aes(Entity, bigoff$`Mismanaged waste (% global total)`)) +
        geom_point() +
        labs(x="Country", y= "Mismanaged waste (% global total)", title= "Most Mismanaged Waste as a % of the Global Total (2010)")
                            
```
##  Conclusion 
This analysis shows that the largest problem is the steep upward trajectory of global plastics production and the world's inability to dispose of it in a sustainable manner.  In the histograms showing global totals of mismanaged waste(% global total), the first histogram shows a high concentration in about 5% of the global total, with a range from 0 to 30%.  We then look at only those countries with an above the mean percentage of the global total. This second histogram shows a range for mismanaged waste from 0 to 5% for this population, with a higher concentration in 0 to 2% of this population, but with predictably higher frequencies across all categories than the first histogram.  

In the second set of histograms plotting per capita plastic waste (kg/person/day) the results show that on a global basis, each person uses between 0.0 to 0.7 kg/person/day, with the largest concentration around 0.0 to 0.3 kg/person/day.  When we again pull out the above the mean users, we see that the per capita plastic waste (kg/person/day) goes from 0.1 to 0.4 kg/person/day, with the highest concentration being from 0.1 to .25 kg/person/day.  

Lastly, we plot the biggest offenders for both per capita plastic and mismanagement to show the top offender for both categories is Sri Lanka. The United States comes in second for plastic waste, and Thailand come second for mismanaged waste.  

